---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 

```{r}
#setting working directory
setwd("~/Desktop/3. Semester/Experimental methods 3/Portfolio-5")

#reading relevant packages
library(metafor)
library(pacman)
library(lmerTest)

#reading the data
data=read.csv("Data.csv", sep=";")

#Effect sizes
Pitch=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = data)

#Making a mixed effects model with study/article as  random effect
Model1 = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=Pitch, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(Model1)

#Meta-analysis optimization
model2 = rma(yi, vi, data = Pitch, slab=Article)
forest(model2)
summary(model2)

#Checking sample sizes 
PitchES=subset(Pitch[Pitch$vi>0,])
sum(PitchES$SampleSizeSchizo, na.rm=TRUE)#166
sum(PitchES$SampleSizeContros, na.rm=TRUE)#95

#Standard Errors
Pitch1=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = data)

#Making a mixed effects model with study/article as a random effect
model3 = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=Pitch1, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(model3)

#Meta-analysis optimization
model4 = rma(yi, vi, data = Pitch1, slab=Article)
forest(model4)
summary(model4)

#Checking sample sizes 
PitchSE=subset(Pitch1[Pitch1$vi>0,])
sum(PitchSE$SampleSizeSchizo, na.rm=TRUE)#257
sum(PitchSE$SampleSizeContros, na.rm=TRUE)#156

#yi = Cohen's D, vi= Variance of Cohen's D
#summary(model) - gives Tau^2 (how much variance between studies), I^2 procentage og tau^2 that could not ... close to 0 if studies variance is overlapping, close to 100 if studies variance is not overlapping at all

```

##Reporting
Effect sizes.
Start with calculating Cohne's D (Yi) and the variance for Cohen's D (Vi). We maked a mixed effect model with study/article as a random effect. 
We made a rma - forest plot to represent this. 
 
4 studies
 participants with 166 Schizophrenia and 95 comparison participants.
 The overall estimated difference (Cohen’s d) in pitch variability between the schizophrenia and the comparison groups was 0.18, SE= 0.17
 
 Standard Errors
 
6 studies
 participants with 257 Schizophrenia and 156 comparison participants.
 The overall estimated difference (Cohen’s d) in pitch variability between the schizophrenia and the comparison groups was -0.125, SE= 0.58

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

```{r}
#nice to extract from model -not necessary
P3Data=read.csv("AllData.csv")
P3Data=P3Data[-1]

PitchMeanSchizo=mean(P3Data$mean [P3Data$Diagnosis=="Schizophrenia"])
PitchMeanControls=mean(P3Data$mean [P3Data$Diagnosis=="Control"])
PitchMeanSchizoSD=sd(P3Data$mean [P3Data$Diagnosis=="Schizophrenia"])
PitchMeanControlsSD=sd(P3Data$mean [P3Data$Diagnosis=="Control"])

PitchSDSchizo=mean(P3Data$SD [P3Data$Diagnosis=="Schizophrenia"])
PitchSDControls=mean(P3Data$SD [P3Data$Diagnosis=="Control"])

PitchSDSchizoSD=sd(P3Data$SD [P3Data$Diagnosis=="Schizophrenia"])
PitchSDControlsSD=sd(P3Data$SD [P3Data$Diagnosis=="Control"])

SampleSizeSchizo=length(unique(P3Data$Subject [P3Data$Diagnosis=="Schizophrenia"]))
#75

SampleSizeContros=length(unique(P3Data$Subject [P3Data$Diagnosis=="Control"]))
#74

OurStudy=data.frame(Article="StudyGroup et. al 2017", Year="2017", SampleSizeSchizo, SampleSizeContros, PitchMeanControls, PitchMeanControlsSD, PitchMeanSchizo, PitchMeanSchizoSD, PitchSDControls, PitchSDControlsSD,PitchSDSchizo, PitchSDSchizoSD)

AllStudies=rbind(data,OurStudy)

#Doing meta-analysis again with new data

PitchAll=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = AllStudies)

#Making a mixed effects model with study/article as  random effect
Model5 = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=PitchAll, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(Model5)

#Meta-analysis optimization
model6 = rma(yi, vi, data = PitchAll, slab=Article)
forest(model6)
summary(model6)

#Checking sample sizes 
PitchES1=subset(PitchAll[PitchAll$vi>0,])
sum(PitchES1$SampleSizeSchizo, na.rm=TRUE) #241
sum(PitchES1$SampleSizeContros, na.rm=TRUE) #169

#Standard Errors
PitchAll1=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = AllStudies)

#Making a mixed effects model with study/article as a random effect
model7 = lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=PitchAll1, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(model7)

#Meta-analysis optimization
model8 = rma(yi, vi, data = PitchAll1, slab=Article)
forest(model8)
summary(model8)

#Checking sample sizes 
PitchSE1=subset(PitchAll1[PitchAll1$vi>0,])
sum(PitchSE1$SampleSizeSchizo, na.rm=TRUE) #332
sum(PitchSE1$SampleSizeContros, na.rm=TRUE) #230

```

Effect sizes.
Start with calculating Cohne's D (Yi) and the variance for Cohen's D (Vi). We maked a mixed effect model with study/article as a random effect. 
We made a rma - forest plot to represent this. 
 
5 studies
 participants with 241 Schizophrenia and 169 comparison participants.
 The overall estimated difference (Cohen’s d) in pitch variability between the schizophrenia and the comparison groups was 0.17, SE= 0.13
 
 Standard Errors
 
7 studies
 participants with 332 Schizophrenia and 230 comparison participants.
 The overall estimated difference (Cohen’s d) in pitch variability between the schizophrenia and the comparison groups was -0.12, SE= 0.43s
 
 
3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

```{r}
summary(model6)
#tau^2=0.0208(SE=0.0552)
#I^2=26.15%

summary(model8)
#tau^2=3.4995(SE=2.0678) 
#I^2=98.18% (very close to 100% - Cohen et al. 2014 - seems not to be from same population - heterogeneity)

#yi = Cohen's D, vi= Variance of Cohen's D
#summary(model) - gives Tau^2 (how much variance between studies), I^2 procentage og tau^2 that could not ... close to 0 if studies variance is overlapping, close to 100 if studies variance is not overlapping at all


#Funnel plots

funnel(model6, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
#the datapoints lie within - but not on both sides - most on the right side (higher than) of the true effect size - evidence for publication bias

regtest(model6)
ranktest(model6)
#Both of them are not significant

funnel(model8, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
#Big outlier - Cohen et al 2014, others lie close together. Not too good. (publication bias?)

regtest(model8)
#Significant funnel plot assymetry. 

ranktest(model8)

inf = influence(model6)
print(inf)
plot(inf)

#Martinez et al. - significant influence 

inf1= influence(model8)
print(inf1)
plot(inf1)

#Cohen et al. 2015 - significant influence

```

## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=0
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2
